Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-13 paper/tv.tex
m

@@

Specification and verification of real-time systems are essential 
to research topics with practical implications. During the last 
more than two decades, a popular approach for specifying real-time systems has been 
based on Timed Automata \cite{DBLP:journals/tcs/AlurD94}. 
Timed Automata are powerful 
in designing real-time models with explicit clock variables. Real-time 
constraints are captured by explicitly setting/resetting clock variables. 
A number of automatic verification support for Timed Automata have proven 
to be successful \cite{DBLP:journals/tse/WangSWQ17,DBLP:journals/sttt/LarsenPY97,yovine1997kronos,wang2005verifying}.

@@

We investigate an alternative approach for modeling and verifying 
compositional real-time systems. In this work, 
we propose a novel temporal specification language, 
which enables a compositional verification via a  Hoare-style 
forward verifier and a term rewriting system (TRS). 
More specifically, we specify system behaviors in the form of 
{Timed Effects} (\timedEffects), which integrates the Kleene Algebra 
with dependent values and arithmetic constraints, 
%gaining the expressive power beyond finite-state machines; and (ii) introduces a new parallel operator 
to 
%capture the parallelism between event-triggered (sync) and  time-triggered (async) execution traces, 
provide real-time abstractions into traditional linear temporal logics. 
%Resource usages, provided by a \emph{dynamic ticks}\cite{von2017real}, incorporating the time-triggered (physical time) and event-triggered (logical time) execution loops. 
For example, one safety property, \textit{"The event \anyevent{Done} 
will be triggered no later than one time unit"}\footnote{Without loss of generality, 
we use integer values to represent time units in this 
paper.% while it can bes extended to real numbers and other 
%time measurement units.
}, is expressed in \timedEffects\ as: 
\begin{align*}
  \code{\effect \defeq \  0 {\leq} t {<}1 \wedge ({\_^\star} \cdot \anyevent{Done} ) \mysharp  t}
\end{align*}  
%The timed effects incorporates different kinds of existing temporal logics, such as linear-time temporal logic (LTL)
%and .
%which replaces temporal operators by time-constrained versions from 
Here, \code{\wedge} connects the arithmetic formula and the timed trace, \code{\mysharp } is a novel operator specifying the \emph{real-time} 
constraints for the \emph{logical-time} sequences \cite{von2017real}; 
\code{\_} is a wildcard matching to any event; 
Kleene star \code{\star} denotes trace repetition.
The above formula \code{\effect} corresponds to `\code{\Diamond_{[0, 1)}\ }\anyevent{Done}' 
in metric temporal logic (MTL), reads \textit{"within one second, 
\anyevent{Done} finally happens"}.

@@

A TRS is a refutation method that normalizes expressions in such a way 
that checking their inclusion corresponds to an iterated process of 
checking the inclusion of their \emph{partial derivatives} 
\cite{antimirov1995partial}. 
Works based on such a TRS \cite{song2020automated,DBLP:journals/tcs/AntimirovM95,almeida2009antimirov,keil2014symbolic,hovland2012inclusion}  show its feasibility and suggest that this method is a better average-case algorithm than those based on the comparison of automata.

@@

The preconidtion of \codem{main} requies no arithmetic constraints 
on its inputs (expressed as \codem{true}) and 
an empty history trace (denoted using \codem{emp} or \code{\epsilon}).
The postcondition of \codem{main} specifies 
that before the final \anyevent{Done} happens, there is no 
occurrence of \anyevent{Done} (\codem{!A} refers to the 
negation of the event \codem{A}). Moreover, the whole process takes 
no more than eight time-units to hit the final event.

\timedEffects\ supports more features such as \emph{disjunctions}, 
\emph{guards} and \emph{parallelism}, etc, which can be found in 
\secref{subsec:Specification_language} and 
\secref{subsec:Specification_Semantics}.  
Note that directly from the specifications, we are aware of:  
(i) the branching properties: different arithmetic conditions on the input parameters lead to different temporal effects;
and (ii) the required history traces: by defining the prior effects 
in precondition. 
The examples 
already show that our  effects provide more detailed information than 
traditional timed verification, and in fact, it cannot be fully 
captured by any prior works \cite{DBLP:journals/tse/WangSWQ17,DBLP:journals/sttt/LarsenPY97,yovine1997kronos,wang2005verifying}.
Nevertheless, the gain in expressive power comes from the efforts 
of a more dedicated verification process, handled by our forward verifier and TRS.

@@

{
\begin{figure}[!ht]
    %  \vspace{-4mm}
     \begin{minipage}[c]{\columnwidth}
         \centering
         {%\small
         \begin{enumerate}
\item \lstinline|void addOneSugar() {| \textcolor{mGray}{\footnotesize 
\quad \emph{// initialize the effects state using the function precondition \code{\effect_{pre}^{addOneSugar(n)}}}} \\
\textcolor{darklavender}{\{\code{\effect_C {=}\effect_{pre}^{addOneSugar(n)} =  true \wedge  \_^\star \} }} 
\quad \siderule{\scriptsize[FV\text{-}Meth] } \\

\item ~\qquad \lstinline|timeout ((), 1);}| \\
\textcolor{darklavender}{\code{\effect_C^\prime {=} t_0{>}1   \wedge \_^\star \cdot \epsilon \mysharp t_0  }} 
\quad \siderule {\scriptsize [FV\text{-}Timeout]}
\item 
\textcolor{darkred}{\code{
%\code{(n{=}0 \wedge \anyevent{Done})\ \vee\ }\code{ (n  \neq 0  \wedge \anyevent{Send}\cdot \effect_{post}^{addNSugar(n-1)}) } 
\effect_C^\prime \CONTAIN \effect_{pre}^{addOneSugar(n)} \cdot \effect_{post}^{addOneSugar(n)} \Leftrightarrow  }} 
 %\textcolor{mGray}{  \emph{(-- check postcondition --) }}
 \textcolor{darkred}{\code{\code{\code{t_0{>}1   \wedge \epsilon \mysharp t_0}
 \CONTAIN t{>}1   \wedge \epsilon \mysharp t }}}
 \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: postcondition checked }}
 \\ 
%ens: t>1 /\ emp#t */ { 
  \noindent\rule{0.92\columnwidth}{0.4pt}
  \vspace{2mm}
~\\
\item \lstinline|void addNSugar (int n) {| \textcolor{mGray}{\footnotesize 
\quad \emph{// initialize the effects state using the function precondition \code{\effect_{pre}^{addNSugar(n)}}.}} \\
     \textcolor{darklavender}{\{\code{\effect_C {=}\effect_{pre}^{addNSugar(n)} =  true \wedge  \_^\star \} }} 
      \quad \siderule{\scriptsize[FV\text{-}Meth] } \\
     \item ~\qquad \lstinline|if (n == 0) {| \\
     \textcolor{darklavender}{\code{\{ n{=}0 \wedge  \_^\star \}  }} 
     \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
      \item ~\qquad~\qquad \lstinline|event ["Sugar"];} |  \\
     \textcolor{darklavender}{\code{\{n{=}0 \wedge   \_^\star \cdot Sugar\} }} 
     \quad \siderule{\scriptsize [FV\text{-}Event]}
     \item 
     ~\qquad \lstinline|else {| \\
     \textcolor{darklavender}{\code{\{ n  {\neq} 0 \wedge   \_^\star \}  }} 
     \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
    \item 
     ~\qquad~\qquad \lstinline|addOneSugar();| \\
     \textcolor{darklavender}{\code{\{n  {\neq} 0 \wedge  t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t  }\}} 
     \quad \siderule {\scriptsize [FV\text{-}Call]}
     \item 
     ~\qquad~\qquad \lstinline|addNSugar (n-1);}}|  \\
     \textcolor{huntergreen}{\code{n  {\neq} 0   \wedge t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t   \CONTAIN \effect_{pre}^{addNSugar(n\text{-}1)} }}
      \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: precondition checked.}}
\\
     \textcolor{darklavender}{\code{\{n  {\neq} 0  \wedge t{>}1  \wedge \_^\star 
     \cdot \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)} }\}} 
     \quad \siderule {\scriptsize [FV\text{-}Call]}
      \item \textcolor{darklavender}{\code{\effect_C^\prime {=} (n{=}0 \wedge  \_^\star \cdot Sugar)\ \vee\ }
      \code{ (n  {\neq} 0  \wedge t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)} ) }} 
      \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
      %\quad \siderule {\scriptsize [Normalization]}
      \\
       \item 
       %\textcolor{mGray}{  \emph{(-TRS: check the  postcondition; Succeed, cf. \autoref{tab:rewriting_tree_send}-) }}\\
       \textcolor{darkred}{\code{
       %\code{(n{=}0 \wedge \anyevent{Done})\ \vee\ }\code{ (n  \neq 0  \wedge \anyevent{Send}\cdot \effect_{post}^{addNSugar(n-1)}) } 
       \effect_C^\prime \CONTAIN \effect_{pre}^{addNSugar(n)} \cdot \effect_{post}^{addNSugar(n)} \Leftrightarrow  }}  
       \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: postcondition checked, cf. \autoref{tab:rewriting_tree_send}. }}
        \\ 
        \textcolor{darkred}{\code{\code{\code{(n{=}0 \wedge Sugar)\ \vee\ }
        \code{ (n  {\neq} 0  \wedge  t{>}1  \wedge \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)}) } \CONTAIN \effect_{post}^{addNSugar(n)}}}}\\

@@

%The rule \code{[REOCCUR]} finds the syntactic identity, as a companion, of the current open goal, as a bud, from the internal proof tree \cite{brotherston2005cyclic}. (We use \code{\textcolor{persianplum}{(\dagger)}} in \tabref{tab:rewriting_tree_send} to indicate such pairings.)

@@

  {
\begin{prooftree}
\Hypo{
  \code{\textcircled{4}\siderule{[\codeme{PROVE}]}}
}

\Infer[dashed]1[]{ \code{\textcolor {darkred}{n{=}0} \wedge \epsilon 
\CONTAIN \textcolor {darkred}{t_r{\geq}0} \wedge \epsilon \mysharp t_r}}

%\textcolor{blue}{\code{\pi_u : true}}
\Infer[dashed]1[{\textcircled{3}\siderule{[\codeme{UNFOLD}]}}]{ \code{ \textcolor {darkred}{n{=}0} \wedge \cancel{Sugar}
\CONTAIN \textcolor {darkred}{t_r{\geq}0} \wedge \cancel{Sugar \mysharp t_r}}}

\Hypo{\qquad \code{(II)}}

\Infer[dashed]2[{\textcircled{2}\siderule{[\codeme{LHS\text{-}OR}]}}]{\code{(\textcolor {darkred}{n{=}0} \wedge Sugar)\ \vee\ }
\code{ (\textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge \epsilon \mysharp t \cdot 
 Sugar \mysharp t_l) } 
\CONTAIN \textcolor {darkred}{t_r{\geq}n} \wedge Sugar \mysharp t_r  }

\Infer[dashed]1[{\textcircled{1}\siderule{[\codeme{RENAME}]}}]{\code{(\textcolor {darkred}{n{=}0} \wedge Sugar)\ \vee\ }
\code{ (\textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1}  \wedge \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)}) } 
\CONTAIN \effect_{post}^{addNSugar(n)}  }
\end{prooftree}}
\\~\\

@@

{\begin{prooftree}
  \Hypo{
  \code{\textcolor {darkred}{  t{>}1 \wedge t_l{\geq}(n\text{-}1) 
   \wedge  \textcolor{blue}{t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1}  \Rightarrow t_r{\geq}n
}  } 
 }

 \Infer[dashed]1[{\textcircled{7}\siderule{[\codeme{PROVE}]}}]{ \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge 
 \epsilon
\CONTAIN \textcolor {darkred}{t_r{\geq}n} 
\wedge \epsilon}}

 \Infer[dashed]1[{\textcircled{6}\siderule{[\codeme{UNFOLD}]}} \textcolor{blue}{\code{\pi_u :  (t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1)}}]
 { \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge 
\cancel{ Sugar \mysharp t_l}
\CONTAIN \textcolor {darkred}{t_r{\geq}n} 
\wedge \cancel{Sugar \mysharp t_r^1}}}

 \Infer[dashed]1[{\textcircled{5}\siderule{[\codeme{UNFOLD}]}} \textcolor{blue}{\code{\pi_u :  (t_r{=}t_r^1 {+}t)}}]
 { \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1) } \wedge \cancel{\epsilon \mysharp t} \cdot 
 Sugar \mysharp t_l
\CONTAIN \textcolor {darkred}{t_r{\geq}n} \wedge Sugar \cancel{\mysharp t_r}}}

@@

In the second branch, step \textcircled{5} eliminates the 
head of the antecedent, where the head is a time duration \code{\epsilon\mysharp t}. 
Therefore the rule $\codeme{[UNFOLD]}$ subtract the time duration 
from the consequent, leaving a fresh time variable \code{t_r^1} and 
an unification constraint \textcolor{blue}{\code{\pi_u : t_r{=}t_r^1 {+}t}}. 
Similarly, step \textcircled{6} eliminates the head 
\code{\anyevent{Sugar}\mysharp t_l}, adding \textcolor{blue}{\code{t_l{=}t_r^1}} 
to the unification constraints. 
At the end of the rewriting, step \textcircled{7} manages to prove that 
\code{\textcolor {darkred}{  t{>}1 \wedge t_l{\geq}(n\text{-}1) 
\wedge  \textcolor{blue}{t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1}  \Rightarrow t_r{\geq}n
}  } \footnote{The proof obligations generated by the verifier are discharged using constraint solver Z3 \cite{de2008z3}.}; therefore, the proof succeed.

%(\code{\{Go\} \Rightarrow \_})
%(\code{\{Ready\} \Rightarrow \_}, \code{t_L^2 \Rightarrow true})
Termination is guaranteed because the set of derivatives to be considered is finite, and possible cycles are detected using 
\emph{memorization} \cite{brotherston2005cyclic}, cf., \tabref{tab:reoccur}.

@@

We show that our method is able to prove the algorithm in the 
section of implementation.

@@

As shown in \figref{fig:Sementic}, we define the semantics of \timedEffects. 
We use 
$[]$ to represent the empty sequence;
$\lappend$ to represent the append operation of two traces;
\code{[I]} to represent the sequence only contains one event.
\code{{\tt{SAT}(\pi)} } indicates the pure \code{\pi} is satisfiable, which is discharged by the constraint solver Z3 \cite{de2008z3}.

@@

Now, the inclusions are disjunction-free formulas. 
Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{keil2014symbolic}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
 initial \emph{heads}, denoted as \code{h},  of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D^{\env}_{h}(\pi \wedge \es)} computes a next-state 
effects after eliminating the head  \code{h} from the current 
effects \code{{\pi} \wedge \es}, w.r.t \code{\env}.

@@

\begin{table}[ht]
\caption{Normalization Axioms for \timedEffects.}
\label{table:normalization}
\setlength{\tabcolsep}{10pt}
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{rl}
\footnotesize

>>>>

\usepackage{xcolor}
\usepackage[thicklines]{cancel}
\renewcommand{\CancelColor}{\color{lightgray}}

@@

Specification and verification of real-time systems are essential 
to research topics with practical implications. During the last 
more than two decades, a popular approach for specifying real-time systems has been 
based on Timed Automata \cite{DBLP:journals/tcs/AlurD94}. 
Timed Automata are powerful 
in designing real-time models with explicit clock variables. Real-time 
constraints are captured by explicitly setting/resetting clock variables. 
A number of automatic verification support for Timed Automata have proven 
to be successful \cite{DBLP:journals/tse/WangSWQ17,DBLP:journals/sttt/LarsenPY97,yovine1997kronos,DBLP:conf/seke/WangWH05}.

@@

We investigate an alternative approach for modeling and verifying 
compositional real-time systems. In this work, 
we propose a novel temporal specification language, 
which enables a compositional verification via a  Hoare-style 
forward verifier and a term rewriting system (TRS). 
More specifically, we specify system behaviors in the form of 
{Timed Effects} (\timedEffects), which integrates the Kleene Algebra 
with dependent values and arithmetic constraints, 
%gaining the expressive power beyond finite-state machines; and (ii) introduces a new parallel operator 
to 
%capture the parallelism between event-triggered (sync) and  time-triggered (async) execution traces, 
provide real-time abstractions into traditional linear temporal logics. 
%Resource usages, provided by a \emph{dynamic ticks}\cite{DBLP:conf/fdl/HanxledenBG17}, incorporating the time-triggered (physical time) and event-triggered (logical time) execution loops. 
For example, one safety property, \textit{"The event \anyevent{Done} 
will be triggered no later than one time unit"}\footnote{Without loss of generality, 
we use integer values to represent time units in this 
paper.% while it can bes extended to real numbers and other 
%time measurement units.
}, is expressed in \timedEffects\ as: 
\begin{align*}
  \code{\effect \defeq \  0 {\leq} t {<}1 \wedge ({\_^\star} \cdot \anyevent{Done} ) \mysharp  t}
\end{align*}  
%The timed effects incorporates different kinds of existing temporal logics, such as linear-time temporal logic (LTL)
%and .
%which replaces temporal operators by time-constrained versions from 
Here, \code{\wedge} connects the arithmetic formula and the timed trace, \code{\mysharp } is a novel operator specifying the \emph{real-time} 
constraints for the \emph{logical-time} sequences \cite{DBLP:conf/fdl/HanxledenBG17}; 
\code{\_} is a wildcard matching to any event; 
Kleene star \code{\star} denotes trace repetition.
The above formula \code{\effect} corresponds to `\code{\Diamond_{[0, 1)}\ }\anyevent{Done}' 
in metric temporal logic (MTL), reads \textit{"within one second, 
\anyevent{Done} finally happens"}.

@@

A TRS is a refutation method that normalizes expressions in such a way 
that checking their inclusion corresponds to an iterated process of 
checking the inclusion of their \emph{partial derivatives} 
\cite{antimirov1995partial}. 
Works based on such a TRS \cite{DBLP:conf/icfem/SongC20,DBLP:journals/tcs/AntimirovM95,almeida2009antimirov,DBLP:conf/fsttcs/KeilT14,hovland2012inclusion}  show its feasibility and suggest that this method is a better average-case algorithm than those based on the comparison of automata.

@@

The precondition of \codem{main} requires no arithmetic constraints 
on its inputs (expressed as \codem{true}) and 
an empty history trace (denoted using \codem{emp} or \code{\epsilon}).
The postcondition of \codem{main} specifies 
that before the final \anyevent{Done} happens, there is no 
occurrence of \anyevent{Done} (\codem{!A} refers to the 
negation of the event \codem{A}). Moreover, the whole process takes 
no more than eight time-units to hit the final event.

\timedEffects\ supports more features such as \emph{disjunctions}, 
\emph{guards} and \emph{parallelism}, etc, which can be found in 
\secref{subsec:Specification_language} and 
\secref{subsec:Specification_Semantics}.  
Note that directly from the specifications, we are aware of:  
(i) the branching properties: different arithmetic conditions on the input parameters lead to different temporal effects;
and (ii) the required history traces: by defining the prior effects 
in precondition. 
The examples 
already show that our  effects provide more detailed information than 
traditional timed verification, and in fact, it cannot be fully 
captured by any prior works \cite{DBLP:journals/tse/WangSWQ17,DBLP:journals/sttt/LarsenPY97,yovine1997kronos,DBLP:conf/seke/WangWH05}.
Nevertheless, the gain in expressive power comes from the efforts 
of a more dedicated verification process, handled by our forward verifier and TRS.

@@

{
\begin{figure}[!ht]
    %  \vspace{-4mm}
     \begin{minipage}[c]{\columnwidth}
         \centering
         {%\small
         \begin{enumerate}
\item \lstinline|void addOneSugar() {| \textcolor{mGray}{\footnotesize 
\quad \emph{// initialize the effects state using the function precondition \code{\effect_{pre}^{addOneSugar(n)}}.}} \\
\textcolor{darklavender}{\{\code{\effect_C {=}\effect_{pre}^{addOneSugar(n)} =  true \wedge  \_^\star \} }} 
\quad \siderule{\scriptsize[FV\text{-}Meth] } \\

\item ~\qquad \lstinline|timeout ((), 1);}| \\
\textcolor{darklavender}{\code{\effect_C^\prime {=} t_0{>}1   \wedge \_^\star \cdot \epsilon \mysharp t_0  }} 
\quad \siderule {\scriptsize [FV\text{-}Timeout]}
\item 
\textcolor{darkred}{\code{
%\code{(n{=}0 \wedge \anyevent{Done})\ \vee\ }\code{ (n  \neq 0  \wedge \anyevent{Send}\cdot \effect_{post}^{addNSugar(n-1)}) } 
\effect_C^\prime \CONTAIN \effect_{pre}^{addOneSugar(n)} \cdot \effect_{post}^{addOneSugar(n)} \Leftrightarrow  }} 
 %\textcolor{mGray}{  \emph{(-- check postcondition --) }}
 \textcolor{darkred}{\code{\code{\code{t_0{>}1   \wedge \epsilon \mysharp t_0}
 \CONTAIN t{>}1   \wedge \epsilon \mysharp t }}}
 \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: postcondition checked. }}
 \\ 
%ens: t>1 /\ emp#t */ { 
  \noindent\rule{0.92\columnwidth}{0.4pt}
  \vspace{2mm}
~\\
\item \lstinline|void addNSugar (int n) {| \textcolor{mGray}{\footnotesize 
\quad \emph{// initialize the effects state using the function precondition \code{\effect_{pre}^{addNSugar(n)}}.}} \\
     \textcolor{darklavender}{\{\code{\effect_C {=}\effect_{pre}^{addNSugar(n)} =  true \wedge  \_^\star \} }} 
      \quad \siderule{\scriptsize[FV\text{-}Meth] } \\
     \item ~\qquad \lstinline|if (n == 0) {| \\
     \textcolor{darklavender}{\code{\{ n{=}0 \wedge  \_^\star \}  }} 
     \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
      \item ~\qquad~\qquad \lstinline|event ["Sugar"];} |  \\
     \textcolor{darklavender}{\code{\{n{=}0 \wedge   \_^\star \cdot Sugar\} }} 
     \quad \siderule{\scriptsize [FV\text{-}Event]}
     \item 
     ~\qquad \lstinline|else {| \\
     \textcolor{darklavender}{\code{\{ n  {\neq} 0 \wedge   \_^\star \}  }} 
     \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
    \item 
     ~\qquad~\qquad \lstinline|addOneSugar();| \\
     \textcolor{darklavender}{\code{\{n  {\neq} 0 \wedge  t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t  }\}} 
     \quad \siderule {\scriptsize [FV\text{-}Call]}
     \item 
     ~\qquad~\qquad \lstinline|addNSugar (n-1);}}|  \\
     \textcolor{huntergreen}{\code{n  {\neq} 0   \wedge t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t   \CONTAIN \effect_{pre}^{addNSugar(n\text{-}1)} }}
      \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: precondition checked.}}
\\
     \textcolor{darklavender}{\code{\{n  {\neq} 0  \wedge t{>}1  \wedge \_^\star 
     \cdot \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)} }\}} 
     \quad \siderule {\scriptsize [FV\text{-}Call]}
      \item \textcolor{darklavender}{\code{\effect_C^\prime {=} (n{=}0 \wedge  \_^\star \cdot Sugar)\ \vee\ }
      \code{ (n  {\neq} 0  \wedge t{>}1  \wedge \_^\star \cdot \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)} ) }} 
      \quad \siderule{\scriptsize [FV\text{-}If\text{-}Else]}
      %\quad \siderule {\scriptsize [Normalization]}
      \\
       \item 
       %\textcolor{mGray}{  \emph{(-TRS: check the  postcondition; Succeed, cf. \autoref{tab:rewriting_tree_send}-) }}\\
       \textcolor{darkred}{\code{
       %\code{(n{=}0 \wedge \anyevent{Done})\ \vee\ }\code{ (n  \neq 0  \wedge \anyevent{Send}\cdot \effect_{post}^{addNSugar(n-1)}) } 
       \effect_C^\prime \CONTAIN \effect_{pre}^{addNSugar(n)} \cdot \effect_{post}^{addNSugar(n)} \Leftrightarrow  }}  
       \quad \textcolor{mGray}{\footnotesize  \emph{// TRS: postcondition checked, cf. \autoref{tab:rewriting_tree_send}. }}
        \\ 
        \textcolor{darkred}{\code{\code{\code{(n{=}0 \wedge Sugar)\ \vee\ }
        \code{ (n  {\neq} 0  \wedge  t{>}1  \wedge \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)}) } \CONTAIN \effect_{post}^{addNSugar(n)}}}}\\

@@

%The rule \code{[REOCCUR]} finds the syntactic identity, as a companion, of the current open goal, as a bud, from the internal proof tree \cite{DBLP:conf/tableaux/Brotherston05}. (We use \code{\textcolor{persianplum}{(\dagger)}} in \tabref{tab:rewriting_tree_send} to indicate such pairings.)

@@

  {
\begin{prooftree}
\hypo{\text{\textcircled{4}}
  \code{\siderule{[\codeme{PROVE}]}}
}

\infer[dashed]1[]{ \code{\textcolor {darkred}{n{=}0} \wedge \epsilon 
\CONTAIN \textcolor {darkred}{t_r{\geq}0} \wedge \epsilon \mysharp t_r}}

%\textcolor{blue}{\code{\pi_u : true}}
\infer[dashed]1[{\textcircled{3}\siderule{[\codeme{UNFOLD}]}}]{ \code{ \textcolor {darkred}{n{=}0} \wedge \cancel{Sugar}
\CONTAIN \textcolor {darkred}{t_r{\geq}0} \wedge \cancel{Sugar \mysharp t_r}}}

\hypo{\qquad \code{(II)}}

\infer[dashed]2[{\textcircled{2}\siderule{[\codeme{LHS\text{-}OR}]}}]{\code{(\textcolor {darkred}{n{=}0} \wedge Sugar)\ \vee\ }
\code{ (\textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge \epsilon \mysharp t \cdot 
 Sugar \mysharp t_l) } 
\CONTAIN \textcolor {darkred}{t_r{\geq}n} \wedge Sugar \mysharp t_r  }

\infer[dashed]1[{\textcircled{1}\siderule{[\codeme{RENAME}]}}]{\code{(\textcolor {darkred}{n{=}0} \wedge Sugar)\ \vee\ }
\code{ (\textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1}  \wedge \epsilon \mysharp t \cdot \effect_{post}^{addNSugar(n\text{-}1)}) } 
\CONTAIN \effect_{post}^{addNSugar(n)}  }
\end{prooftree}}
\\~\\

@@

{\begin{prooftree}
  \hypo{
  \code{\textcolor {darkred}{  t{>}1 \wedge t_l{\geq}(n\text{-}1) 
   \wedge  \textcolor{blue}{t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1}  \Rightarrow t_r{\geq}n
}  } 
 }

 \infer[dashed]1[{\textcircled{7}\siderule{[\codeme{PROVE}]}}]{ \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge 
 \epsilon
\CONTAIN \textcolor {darkred}{t_r{\geq}n} 
\wedge \epsilon}}

 \infer[dashed]1[{\textcircled{6}\siderule{[\codeme{UNFOLD}]}} \textcolor{blue}{\code{\pi_u :  (t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1)}}]
 { \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1)}  \wedge 
\cancel{ Sugar \mysharp t_l}
\CONTAIN \textcolor {darkred}{t_r{\geq}n} 
\wedge \cancel{Sugar \mysharp t_r^1}}}

 \infer[dashed]1[{\textcircled{5}\siderule{[\codeme{UNFOLD}]}} \textcolor{blue}{\code{\pi_u :  (t_r{=}t_r^1 {+}t)}}]
 { \code{ \textcolor {darkred}{n  {\neq} 0  \wedge  t{>}1 \wedge t_l{\geq}(n\text{-}1) } \wedge \cancel{\epsilon \mysharp t} \cdot 
 Sugar \mysharp t_l
\CONTAIN \textcolor {darkred}{t_r{\geq}n} \wedge Sugar \cancel{\mysharp t_r}}}

@@

In the second branch, step \textcircled{5} eliminates the 
head of the antecedent, where the head is a time duration \code{\epsilon\mysharp t}. 
Therefore the rule $\codeme{[UNFOLD]}$ subtract the time duration 
from the consequent, leaving a fresh time variable \code{t_r^1} and 
an unification constraint \textcolor{blue}{\code{\pi_u : t_r{=}t_r^1 {+}t}}. 
Similarly, step \textcircled{6} eliminates the head 
\code{\anyevent{Sugar}\mysharp t_l}, adding \textcolor{blue}{\code{t_l{=}t_r^1}} 
to the unification constraints. 
At the end of the rewriting, step \textcircled{7} manages to prove that 
\code{\textcolor {darkred}{  t{>}1 \wedge t_l{\geq}(n\text{-}1) 
\wedge  \textcolor{blue}{t_r{=}t_r^1 {+}t \wedge  t_l{=}t_r^1}  \Rightarrow t_r{\geq}n
}  } \footnote{The proof obligations generated by the verifier are discharged using constraint solver Z3 \cite{DBLP:conf/tacas/MouraB08}.}; therefore, the proof succeed.

%(\code{\{Go\} \Rightarrow \_})
%(\code{\{Ready\} \Rightarrow \_}, \code{t_L^2 \Rightarrow true})
Termination is guaranteed because the set of derivatives to be considered is finite, and possible cycles are detected using 
\emph{memorization} \cite{DBLP:conf/tableaux/Brotherston05}, cf., \tabref{tab:reoccur}.

@@

We show that our method is able to prove the algorithm in the 
section of implementation.  prove the algorithm in the 
section of implementation.

@@

As shown in \figref{fig:Sementic}, we define the semantics of \timedEffects. 
We use 
$[]$ to represent the empty sequence;
$\lappend$ to represent the append operation of two traces;
\code{[I]} to represent the sequence only contains one event.
\code{{\tt{SAT}(\pi)} } indicates the pure \code{\pi} is satisfiable, which is discharged by the constraint solver Z3 \cite{DBLP:conf/tacas/MouraB08}.

@@

Now, the inclusions are disjunction-free formulas. 
Next we provide the definitions and the key implementations\footnote{As 
the implementations according to basic regular expressions can be found 
in prior work \cite{DBLP:conf/fsttcs/KeilT14}. 
%Here, we focus on presenting 
%the definitions and how do we deal with dependent values in the effects, 
%as the key novelties of this work. 
%More detail implementations can be found in the extended technical report\citeref{Demo}.
} of Nullable, First and Derivative respectively. Intuitively, 
the Nullable function \code{\delta_{{\pi}}(\es)} returns a boolean 
value indicating whether \code{{\pi} \wedge \es} contains the empty 
trace; the First function \code{fst_{{\pi}}( \es)} computes a set of 
 initial \emph{heads}, denoted as \code{h},  of \code{{\pi} \wedge \es}; and the Derivative 
function \code{D^{\env}_{h}(\pi \wedge \es)} computes a next-state 
effects after eliminating the head  \code{h} from the current 
effects \code{{\pi} \wedge \es}, w.r.t \code{\env}.

@@

\begin{table}[ht]
\caption{Normalization Axioms for \timedEffects.}
\label{table:normalization}
\setlength{\tabcolsep}{8pt}
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{rl}
\footnotesize
